---
title: "Anthony Fantano Ratings Pre-processing and EDA"
output: html_document
---

The following document takes a deeper look at famous YouTube music reviewer Anthony Fantano's album ratings. The data contains all reviews/scores of music albums rated by him, and was webscraped from rateyourmusic.com/~theneedledrop. _Analyzing his ratings on hip hop albums was an area of interest_

__Data Source:__

* fantano.csv

__Below is a list of variables that were of importance:__

* release_date - the date the album was released.
* fantano_rating - Anthony Fantano’s personal rating of the album (out of 100).
* album_name - the name of the album.
* artist - the name of the artist.
* genre - the genre of the album.
* total_length - the length of the album.
* critic_score - the average score of all other critic scores for the album.
* user_score - the average user score for all users who rated the album on the website.
* producer - the producer(s) of the album.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

#### Reading in data

```{r}
fantano <- read_csv("fantano.csv")
```

#### Filtering out NA values in total_length

```{r}
fantano_filter <- fantano %>%
  drop_na(total_length)
```

#### Creating new column that only contains only numeric value of total_length and with no white space

```{r}
fantano_filter <- fantano_filter %>%
  mutate(album_length = str_trim(str_to_lower(str_replace(fantano_filter$total_length, 'Total Length: ', ''))))
```

#### Creating new column that only contains minute value of album_length and converting those values to doubles

```{r}
fantano_filter <- fantano_filter %>%
  tidyr::extract(album_length, 'album_minutes_only', regex = '(\\d{1,} minute)', remove = F) %>%
  mutate(album_minutes_only = as.double(str_replace(album_minutes_only, 'minute', '')))
```

#### Creating new column that only contains hour value of album_length and creating new column that converts it to minutes

```{r}
fantano_filter<-fantano_filter%>%
  tidyr::extract(col=album_length,into = "hours", regex = "(\\d\\shour)", remove=F)%>%
  tidyr::extract(col=hours, into="hours", regex="(\\d)", remove=F)%>%
  mutate(hours_minutes=as.double(hours))%>%
  mutate(hours_minutes=hours_minutes*60)
```

#### Creating new column that sums hours_minutes and album_minutes_only to find total length of each album in minutes

```{r}
fantano_filter <- fantano_filter %>%
  rowwise() %>%
  mutate(total_length_minutes = sum(c(hours_minutes, album_minutes_only), na.rm = T))
```

#### Finding the album that had the longest length in minutes

```{r}
longest_album <- fantano_filter %>%
  arrange(desc(total_length_minutes)) %>%
  head(1) %>%
  pull(album)

```
```{r, echo=FALSE}
show(longest_album)
```


#### Creating a new column that is set to 1 if 'hip hop' is detected as a genre

```{r}
fantano <- fantano %>%
  mutate(hip_hop = ifelse(str_detect(tolower(genres), "hip hop"), 1, 0))
```

#### Creating a new column that finds the number of producers of each album

```{r}
fantano <- fantano %>%
  mutate(producer_number = stringr::str_count(producer, ",") + 1 )
```

#### Finding the hip hop album with the most amount of producers

```{r}
most_producers <- fantano %>%
  arrange(desc(producer_number)) %>%
  head(1) %>%
  pull(album)
```
```{r, echo=FALSE}
show(most_producers)
```


#### Creating a new column in Fantano that is the release_date column in YYYY-MM-DD format, filtering to only include hip hop albums released in 2022. 

#### Creating a new tibble that only contains the artist, album, and fantano_rating columns from Fantano data set

```{r}
fantano <- fantano %>%
  mutate(date = lubridate::mdy(release_date)) %>%
  filter(str_detect(date, "2022")) %>%
  filter(hip_hop == 1)

fantano_graph <- fantano %>%
  select(artist, album, fantano_rating)
```

#### Loading in glue library and creating a new column that contains the artist and album name

```{r}
library(glue)
fantano_graph <- fantano_graph %>%
  mutate(artist_album = glue('{fantano_graph$artist}: {fantano_graph$album}'))
```

#### Finding the top 15 highest rated hip hop albums by Fantano

```{r}
fantano_graph_subset <- fantano_graph %>%
  arrange(desc(fantano_rating)) %>%
  head(15)
```
```{r, echo=FALSE}
show(fantano_graph_subset)
```


#### Visualizing the top 15 hip hop albums of 2022 according to Fantano

```{r}
fantano_graph_final <- fantano_graph_subset %>%
  mutate(artist_album = fct_reorder(artist_album, fantano_rating)) %>%
  ggplot(aes(x = fantano_rating, y = artist_album)) + 
  geom_col() + 
  coord_cartesian(xlim = c(60,90)) + 
  labs(x = "Fantano Score",
       y = "",
       title = "Top 15 Hip Hop Albums of 2022",
       subtitle = "Anthony Fantano Ratings") + 
  theme_minimal()
```
```{r, echo=FALSE}
plot(fantano_graph_final)
```

#### Reloading fantano data

```{r}
fantano_writeup <- read_csv('fantano.csv')
popular_genres <- c('Hip Hop', 'Indie Rock', 'Trap', 'Southern Hip Hop', 'Alternative Rock', 'Electronic')

```

#### Finding the average difference between fantano_rating and critic_score for the top 10 highest rated fantano albums

```{r}
highest_fantano_ratings <- fantano_writeup %>%
  arrange(desc(fantano_rating)) %>%
  select(album, fantano_rating, critic_score) %>%
  mutate(difference = fantano_rating - critic_score) %>%
  head(10) %>%
  mutate(average_difference = mean(difference))
```
```{r, echo=FALSE}
show(highest_fantano_ratings)
```


#### Finding the average difference between fantano_rating and critic_score for the top 10 lowest rated fantano albums

```{r}
lowest_fantano_ratings <- fantano_writeup %>%
  drop_na() %>%
  arrange(desc(fantano_rating)) %>%
  select(album, fantano_rating, critic_score) %>%
  mutate(difference = fantano_rating - critic_score) %>%
  tail(10) %>%
  mutate(average_difference = mean(difference))
```
```{r, echo=FALSE}
show(lowest_fantano_ratings)
```

#### Visualizing the difference between the top 10 highest fantno ratings and their critic equivalent rating

```{r}
top_ranking_difference <- highest_fantano_ratings %>%
  ggplot(aes(y = album)) +
  geom_col(aes(x = fantano_rating, fill = 'coral2')) +
  geom_col(aes(x= critic_score, fill = 'cornflowerblue'))+
  scale_fill_manual(
    name = 'Reviewer',
    values = c('coral2' = 'coral2', 'cornflowerblue' =
                 'cornflowerblue'),
    labels = c('Fantanto Score', 'Critic Score')
  ) +
  labs( x = 'Rating',
        y = 'Album',
        title = 'Top 10 Highest Fantano Ratings and Their Critic Equivalent')
```
```{r, echo=FALSE}
plot(top_ranking_difference)
```

#### Visualizing the difference between the lowest 10 fantno ratings and their critic equivalent rating

```{r}
lowest_ranking_difference <- lowest_fantano_ratings %>%
  ggplot(aes(y = album)) +
  geom_col(aes(x= critic_score, fill = 'cornflowerblue'))+
  geom_col(aes(x = fantano_rating, fill = 'coral2')) +
  scale_fill_manual(
    name = 'Reviewer',
    values = c('coral2' = 'coral2', 'cornflowerblue' =
                 'cornflowerblue'),
    labels = c('Fantanto Score', 'Critic Score')
  ) +
  labs( x = 'Rating',
        y = 'Album',
        title = 'Top 10 Lowest Fantano Ratings and Their Critic Equivalent')
```
```{r, echo=FALSE}
plot(lowest_ranking_difference)
```

#### Finding average fantano, critic, and user scores

#### Finding the difference between Fantano's ratings and the critic's ratings and also Fantano's ratings and the user's ratings

```{r}
average_fantano_rating <- mean(fantano_writeup$fantano_rating, na.rm = T)
average_critic_rating <- mean(fantano_writeup$critic_score, na.rm = T)
average_user_rating <- mean(as.double(fantano_writeup$user_score), na.rm = T)
average_critic_and_user <- (average_critic_rating + average_user_rating) /2
difference_in_critic_rating <- average_fantano_rating - average_critic_rating
difference_in_user_rating<- average_fantano_rating - average_user_rating
```
```{r, echo=FALSE}
show(average_fantano_rating)
show(average_critic_rating)
show(average_user_rating)
show(average_critic_and_user)
show(difference_in_critic_rating)
show(difference_in_user_rating)
```















